---
title: "p8105_hw1_xf2290"
author: "Xiwei Fu"
output: github_document
date: "2025-09-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Problem 1
```{r}
library(moderndive)
library(ggplot2) 
library(tidyverse)#load packages
data("early_january_weather") #load dataset
```
Short description of the dataset
This dataset is hourly meteorological data for LGA, JFK and EWR for the month of January 2013, which includes 358 rows representing hourly measurements and 15 variables.
The variables are:
origin, year, month, day, hour, temp, dewp, humid, wind_dir, wind_speed, wind_gust, precip, pressure, visib, time_hour.

```{r}
mean(early_january_weather %>% pull(temp))
```
We have the mean temperature is 39.58212 F.
```{r}
ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point() +
  labs(
    title = "Temperature vs Time in Early January 2013",
    x = "Time (hourly observations)",
    y = "Temperature (°F)",
    color = "Humidity (%)"
  ) 
ggsave("temp_plot.png")
```

Problem 2
```{r}
df <- tibble(
  norm_sample = rnorm(10),                                  # numeric
  if_positive = rnorm(10) > 0,                              # logical vector
  char_vec  = sample(letters[1:3], 10, replace=TRUE),     
# character vector
  fact_vec = factor(sample(c("low","med","high"), 10, replace=TRUE)) # factor vector
)
```
```{r}
# Numeric variable
mean(df %>% pull(norm_sample))

# Logical variable
mean(df %>% pull(if_positive))

# Character variable
mean(df %>% pull(char_vec))

# Factor variable
mean(df %>% pull(fact_vec))
```
For Character variables, it fails, because we cannot average the characters.
For Factor variable, it fails, because we need conversion before calculate it.
For Numeric variable, it works.
For Logical variable, it works, because 'true' is treated as 1 and 'false' is treated as 0. 
```{r}
as.numeric(df %>% pull(if_positive))
as.numeric(df %>% pull(char_vec))
as.numeric(df %>% pull(fact_vec))
```
This help explain what happens when I try to take the mean.
For Logical variable, it shows me 0 and 1.
For Character variable, it shows NA since letters cannot be coerced to numbers.
For Factor variable, it returns the integer codes for the factor levels, it means we can calculate the mean for it after conversion.
